Interisland movements and oceanic swimming of woodland caribou in Newfoundland

### 2020-05-11 paper/manuscript.Rmd
merge p4, p5
trade-offs

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, the energetic expenditure
of swimming for quadrupedal mammals is significantly higher than walking or
running, and drowning is also possible [@Miller1985]. Despite
the abundance of coastal and island caribou herds, only a handful of examples
have documented caribou swimming in the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012].

Caribou may swim between islands in the ocean for similar reasons they swim in
freshwater [@Leblond2016], mainly concerning predator avoidance or movement
between habitats. We propose that one such explanation, the forage limitation
hypothesis, is the most likely explanation for oceanic swimming for caribou that
live on islands in the Fogo Island archipelago.

@@

@@

Movement between islands requires caribou to assess the trade-off associated
with swimming. This trade-off can be understood as density-dependent habitat
selection governed by the Ideal Free Distribution [@Morris1987; @Bradbury2015].
Ideal Free Distribution theory predicts a fitness equilibrium: when the density
in a given habitat patch has exceeded the optimum for fitness within that patch,
animals should relocate and settle new habitat patches, so that fitness is equal
across all patches [@Bradbury2015]. Islands act as discrete habitat patches in
this case, but the costs associated with swimming create a trade-off when moving
between patches. For caribou in the Fogo Island archipelago, the role of forage
limitation in habitat selection patterns remains unknown, but swimming
between islands may be at least partially governed by density-dependent habitat selection.

>>>>

Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond2016],
to avoid predators [@Bergerud1985], and access islands during calving
[@Bergerud1990]. Even for adept swimmers like caribou, the energetic expenditure
of swimming for quadrupedal mammals is significantly higher than walking or
running, and drowning is also possible [@Miller1985]. Despite the abundance of
coastal and island caribou herds, only a handful of examples have documented
caribou swimming in the ocean [e.g. @Miller1995; @Miller2002;
@Jeffery2007; @Ricca2012]. Caribou may swim between islands in the ocean for
similar reasons they swim in freshwater [@Leblond2016], mainly concerning
predator avoidance or movement between habitats. We propose that one such
explanation, the forage limitation hypothesis, is the most likely explanation
for oceanic swimming for caribou that live on islands in the Fogo Island
archipelago.

@@

The trade-off associated with swimming between islands can be understood as
density-dependent habitat selection governed by the Ideal Free Distribution
[@Morris1987; @Bradbury2015]. Ideal Free Distribution theory predicts a fitness
equilibrium: when the density in a given habitat patch has exceeded the optimum
for fitness within that patch, animals should relocate and settle new habitat
patches, so that fitness is equal across all patches [@Bradbury2015]. Islands
act as discrete habitat patches in this case, but the costs associated with
swimming create a trade-off when moving between patches. For caribou in the Fogo
Island archipelago, the role of forage limitation in habitat selection patterns
remains unknown, but swimming between islands may be at least partially governed
by density-dependent habitat selection.

@@
